# Some Python Basics

```{r echo=F}
my_section <- function(title){
  s <- paste('## ',title,' {}')
  return(s)
}

```

`r my_section('Python is Case Sensitive')`

This will work.

```{r echo = T, engine = 'python'}
print('Hello World')
```

This will **not**.

```{r echo = T, engine = 'python', error=T}
Print('Hello World')
```

`r my_section('<code>x</code> is a variable. \'x\' is an English letter')`

We asking Python to print the English letter 'x'.
```{r echo = T, engine = 'python'}
print('x')
```

We asking Python to print the value related to the **variable** `x`.
```{r echo = T, engine = 'python'}
x = 10
print(x)
```

`r my_section('We can use f\'letters{x}\' to combine English with variables')`

```{r echo = T, engine = 'python'}
x = 10
text = 'My age is {x}.'
print(text)
```

`r my_section('Be mindful of spaces(indentation) at the start of a line')`

This will work.

```{r echo = T, engine = 'python'}
x = 10
y = 20
```

This will **not**.

```{r echo=T, engine='python', eval = F}
x = 10
  y = 20
```


`r my_section('Python has functions')`

`print()` is a good example of a function in Python. For this function to work it needs something (which we call an argument) to print. We need supply this argument within the function's brackets `()`.

```{r echo = T, engine = 'python'}
print('Hello World')
```

A typical function usually can accept additional arguments to do more things. For example the `print()` function allows us to add something to the end.

```{r echo = T, engine = 'python'}
print('Hello World', end = '******')
```

You can find out all about a function my looking at its help file as follows

```{r echo = T, engine = 'python', eval = F}
print?
```
```
print(...)
    print(value, ..., sep=' ', end='\n', file=sys.stdout, flush=False)

    Prints the values to a stream, or to sys.stdout by default.
    Optional keyword arguments:
    file:  a file-like object (stream); defaults to the current sys.stdout.
    sep:   string inserted between values, default a space.
    end:   string appended after the last value, default a newline.
    flush: whether to forcibly flush the stream.
```

`r my_section('Brackets in Python')`

```{r echo=T,engine='python'}
print('Hello World')

```

`r my_section('Python has many (many) cool packages')`

```{r echo=T,engine='python'}
print('Hello World')
```
-   Example: `math.sqrt()` means `sqrt()` belonging to `math`

`r my_section('You can think of the dot(\'.\') as indicating ownership')`

```{r echo=T,engine='python'}
print('Hello World')

```

`r my_section('We can store data using lists')`

`r my_section('We can store data using dictionaries')`

`r my_section('We can store data using numpy arrays')`

`r my_section('Python sometimes act like a prima donna when you make an error')`

```{r echo=T,engine='python'}
print('Hello World')

```
