# Using Pandas to Explore COVID-19 Data

```{r}
library(reticulate)
```

## Plans for this section

In the previous two sections, we introduced basics to plotting using `matplotlib` and generating exploratory analysis & visualisations using `pandas.` In this section, we are going to put these skills into practice and explore the COVID19 dataset from GitHub (https://github.com/datasets/covid-19) that is maintained by Johns Hopkins University.

## Housekeeping

It is always good to get a rough idea about a dataset, right at the start.

<!-- Exercise -->
`r my_fun.exercise('What are we dealing with?')`

#### Tasks {-}

1. Load the data from https://raw.githubusercontent.com/datasets/covid-19/main/data/countries-aggregated.csv
1. What is the shape of the data? I.e. how many rows and columns are there?
1. Are there any missing numbers?
1. What are column names?

#### Solution {-}
```{python,echo=F}
import pandas as pd
import matplotlib.pyplot as plt
```

```{python,include=T}
data_url = 'https://raw.githubusercontent.com/datasets/covid-19/main/data/countries-aggregated.csv'
df = pd.read_csv(data_url)

df.shape

df.columns

df.head()

df.describe()

df.info()


```

```{r, echo = F, eval=FALSE}
knitr::kable(head(py$df))
```



<!-- Exercise -->
`r my_fun.exercise('Drop those unnecessary columns?')`

#### Tasks {-}
<br>

1. What are all the columns in the this dataset.
1. Drop unnecessary columns.

#### Solution {-}


This solution specifies what columns to **drop**.

#### Another Solution {-}

This solution specifies what columns to **keep**.
